---
title: "ABC_Bembix_widepriors"
author: ""
date: "March 2021"
output: html_document
---

 
```{r, message=FALSE, warning=FALSE}
library(abc)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(gridExtra)
library(data.table)
library(psych)

source("H_modelselectionABC.R") #file with custom made ABC analysis (for doing the weighted model selection)

#loading the data: summary stats, params scenarios, field data summary stats
##summary stats
getwd()

multimerge = function(path, pattern){
  filenames=list.files(path=path, pattern=pattern, full.names=T)
  rbindlist(lapply(filenames, function(x){read.table(x, header=T)}))
}

sumstats <- multimerge(path="./Data outputs/Wide priors", pattern="Summary_stats*")
nrow(sumstats)

colnames(sumstats) <- c("scenario", "file title", 'RK2', 'RK5', 'RK10', 'RK15', 'RK20', 'RK30', 'RK40',
                      "T0", "T2", "T4",  "T6", "T8", "T10", "T12", "T14",
                      "NA_internal_loops", "NA_all_loops","NA_dens_undirected", "NA_dens_directed",
                      "NA_reciproc","NA_transitivity_und","NA_transitivity_dir")
sumstats <- sumstats %>% select(-T0, -T2, -T4, -T6, -T8, -T10, -T12, -T14)

#drop the NA's, which stems from some runs that don't have interchange between clusters: these are counted as part of the ABC, but would be dropped immediately, because they don't have interchange between clusters.
#sumstats <- sumstats %>% drop_na()
ind_na <- which(is.na(sumstats), arr.ind=TRUE)[,1]
#get sumstats only for non-NA
sumstats <- sumstats %>% slice(-ind_na)
nrow(sumstats)

#Take relative Ripley's K to complete spatial randomness (CRS) (K-CRS)/CRS; CRS=pi*r^2
sumstats <- sumstats %>% mutate(RK2rel = ( RK2 - pi*2^2 )/( pi*2^2 ),
                                RK5rel = ( RK5 - pi*5^2 )/( pi*5^2 ),
                                RK10rel = ( RK10 - pi*10^2 )/( pi*10^2 ),
                                RK15rel = ( RK15 - pi*15^2 )/( pi*15^2 ),
                                RK20rel = ( RK20 - pi*20^2 )/( pi*20^2 ),
                                RK30rel = ( RK30 - pi*30^2 )/( pi*30^2 ),
                                RK40rel = ( RK40 - pi*40^2 )/( pi*40^2 ))

#parameters
params <- multimerge(path="./Data outputs/Wide priors", pattern="Param*")
nrow(params)

head(params)
summary(params)

#values for NA's
params_na <- params %>% slice(ind_na) #--> small values of sigma_lsf mainly

#get only params for non-NA
params <- params %>% slice(-ind_na)
```


###Field data
Master thesis data  
```{r}
##field data summary stats
ss_field <- read.table("./Field data analyses/Summary_stats_field.txt", header=T)
ss_field$X0 <- NULL
ss_field$scenario <- NULL
ss_field[] <- lapply(ss_field[], as.numeric)
colnames(ss_field) <- c("mean_RK", "sd_RK", 'RK2', 'RK5', 'RK10', 'RK15', 'RK20', 'RK30', 'RK40',
                           "NA_internal_loops", "NA_all_loops","NA_dens_undirected", "NA_dens_directed",
                           "NA_reciproc","NA_transitivity_und","NA_transitivity_dir")
ss_field$mean_RK <- NULL
ss_field$sd_RK <- NULL

#Take relative Ripley's K to complete spatial randomness (CRS) (K-CRS)/CRS; CRS=pi*r^2
ss_field <- ss_field %>% mutate(RK2rel = ( RK2 - pi*2^2 )/( pi*2^2 ),
                                RK5rel = ( RK5 - pi*5^2 )/( pi*5^2 ),
                                RK10rel = ( RK10 - pi*10^2 )/( pi*10^2 ),
                                RK15rel = ( RK15 - pi*15^2 )/( pi*15^2 ),
                                RK20rel = ( RK20 - pi*20^2 )/( pi*20^2 ),
                                RK30rel = ( RK30 - pi*30^2 )/( pi*30^2 ),
                                RK40rel = ( RK40 - pi*40^2 )/( pi*40^2 ))

#make a 1D vector of it
ss_field_v <- as.numeric(ss_field[1,])
```


###Indexes
A vector has to be created with the names in of the models corresponding to the different runs.  

```{r}
#keep summary stats with scenario attached to df
sumstats_wscen <- sumstats
#make sure the summary stats are numeric values (scenario will be replaced with 1, 2, 3)
sumstats[] <- lapply(sumstats[], as.numeric) #make sure all are numeric values

sumstats$scenario <- NULL
sumstats$`file title` <- NULL
```


####Correlation between summary stats
Within the Ripley's K, there is high correlation. But less for the high values. We should select distances that are further apart from each other. F.i. RK2, RK10, RK20, RK30, RK40.  
Within Network metrics, there is some correlation. Especially for the same metric for undirected and directed networks. These will be dropped.  
For the network metrics and Ripley's K, there is some structure. But this is expected with such high level network metrics. Variation seems to be high enough.  

```{r}
#make selection
sumstats_sel <- sumstats %>% select("RK2rel","RK5rel", "RK10rel", "RK20rel", "RK30rel", "RK40rel",
                                 "NA_internal_loops", "NA_all_loops", "NA_dens_directed",
                                 "NA_reciproc", "NA_transitivity_dir")

ss_field_sel <- ss_field %>% select("RK2rel","RK5rel", "RK10rel", "RK20rel", "RK30rel", "RK40rel",
                                 "NA_internal_loops", "NA_all_loops", "NA_dens_directed",
                                 "NA_reciproc", "NA_transitivity_dir")

#standardise data, according to simulations
sumstats_sel_sd <- scale(sumstats_sel, center=summarise_all(sumstats_sel, mean), scale=summarise_all(sumstats_sel, sd))
#also standardise field data, according to simulaitons
ss_field_sel_sd <- scale(ss_field_sel[1,], center=summarise_all(sumstats_sel, mean), scale=summarise_all(sumstats_sel, sd))

#Within Ripley's K
pairs(sumstats_sel_sd[sample(nrow(sumstats_sel_sd), 1000),c(1:6)])
#within network metrics
pairs(sumstats_sel_sd[sample(nrow(sumstats_sel_sd), 1000), c(7:11)])

#make field data vector
ss_field_sel_v <- as.numeric(ss_field_sel_sd[1,])
#make a sumstats selected with scenario
sumstats_sel_scen <- sumstats_sel #not with standardised data
sumstats_sel_scen <- bind_cols(sumstats_sel_scen, "scenario"=sumstats_wscen$scenario)
head(sumstats_sel_scen)
```

```{r}
#indexes -> which row corresponds with which model
indexes <- as.vector(unlist(sumstats_wscen[,1]))
str(indexes)
```

####Making Pearson's correlation matrix

```{r}
#calculate pearson's correlation coefficients for all
pcor <- cor(sumstats_sel_sd)
#View(pcor)

n_st <- nrow(pcor)

#take the mean of the pearson's correlation coefficients per variable
#then, mutate to get mean without the 1 from the diagonal of the matrix (which is x*n/(n-1) + 1/n)
mean_pcor <- as.data.frame(pcor) %>% mutate_all(abs) %>% summarise_all(mean) %>% mutate_all(function(x) {x*n_st/(n_st-1) - 1/n_st})
weights <- mean_pcor %>% mutate_all(function(x) {1-x}) #maybe to the power of 2?

weights
```



##Model selection
'postpr' function from the package abc is used to look which model is mostly selected (when summary stats lay closest to field data sumstats). 

```{r}
# modsel_bembix <- postpr(target=ss_field, index=indexes, sumstat=ss_all,tol=0.001,method="rejection")
# summary(modsel_bembix)

modsel_1000 <- create.abcSel(target=ss_field_sel_v, indexes=indexes,
                             results=as.matrix(sumstats_sel_sd), number.accept = 1000,
                             weights=as.numeric(weights[1,]))
summary(modsel_1000)

modsel_100 <- create.abcSel(target=ss_field_sel_v, index=indexes,
                            results=as.matrix(sumstats_sel_sd),
                            number.accept = 100, weights=as.numeric(weights[1,]))
summary(modsel_100)

#source("D:/fbatslee/OneDrive - UGent/Conferences/BES/ABCWorkshop/Code/R/ModelSelection.R")
```

##Summary stats of models that were accepted

Here the priors (transparant red) of the summary stats are plotted and the posterios (black). The green line is the field data.

```{r, message=FALSE, warning=FALSE}
#index_ss <- as.numeric(rownames(modsel_bembix$ss))
sumstats_sel$accepted <- modsel_100$accepted
index_ss <- as.numeric(which(sumstats_sel$accepted==T))
stats_ss <- sumstats_sel[index_ss,]

ggplot(data=gather(stats_ss), aes(x=value)) +
  geom_histogram(aes(x=value, y=..count../sum(..count..))) + facet_wrap(~key, scales='free') +
  geom_vline(data=gather(ss_field_sel[1,]), aes(xintercept=value), color='green', linetype="dashed", size=3)+
  theme(text = element_text(size=12))+
  geom_histogram(data=gather(sumstats_sel), aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5)
```

###Plot of the accepted summary stats with field data

Ripley's K: spatial clustering is quite well formed.  
Network analysis are produced quite well.  
```{r}
#get ripley's K and network analysis values for the accepted models
stats_ss_w_scen <- sumstats_sel_scen[index_ss,]
RKs <- stats_ss_w_scen %>% select("scenario","RK2rel", "RK5rel", "RK10rel", "RK20rel","RK30rel", "RK40rel")
NAs <- stats_ss_w_scen %>% select("scenario", "NA_internal_loops", "NA_all_loops",
                                  "NA_dens_directed","NA_reciproc", "NA_transitivity_dir")

#get field summary stats
field_RK <- ss_field_sel %>%
  select("RK2rel","RK5rel", "RK10rel", "RK20rel","RK30rel", "RK40rel") %>%
  filter(row_number()==1)
field_RK$scenario <- 'field data'

field_NA <- ss_field_sel %>% select("NA_internal_loops", "NA_all_loops",
                                  "NA_dens_directed","NA_reciproc", "NA_transitivity_dir") %>%
  filter(row_number()==1)

#priors

priors_RK <- sumstats_sel_scen  %>%
   select("scenario", "RK2rel","RK5rel", "RK10rel", "RK20rel", "RK30rel", "RK40rel")
priors_NA <- sumstats_sel_scen %>% select("scenario","NA_internal_loops", "NA_all_loops",
                                  "NA_dens_directed","NA_reciproc", "NA_transitivity_dir")

#change colnames to easily plot the functions correctly
colnames(RKs) <- c("scenario",'2', '5', '10', '20', '30', '40')
colnames(field_RK) <- c('2', '5', '10', '20', '30', '40',"scenario")
colnames(priors_RK) <- c('scenario','2', '5', '10', '20','30', '40')


colnames(NAs) <- c("scenario",'internal loops','all loops','density',
                   'reciprocity', 'transitivity')
colnames(field_NA) <- c('internal loops','all loops','density',
                   'reciprocity', 'transitivity')
colnames(priors_NA) <- c("scenario",'internal loops','all loops','density',
                   'reciprocity', 'transitivity')

#Ripley's K graphs
# RKs_tidy <- RKs %>%
#   gather(distance, value, '2':'40') %>%
#   mutate(distance = as.factor(as.numeric(distance))) %>%
#   group_by(scenario, distance)
# RKs_summary <- RKs_tidy %>% summarise(mean_RK = mean(value), sd_RK = sd(value))
# 
field_RK_tidy <- field_RK %>% gather(distance, value, '2':'40') %>%
  mutate(distance = as.factor(as.numeric(distance)))
RKs_tidy <- RKs %>% gather("2","5", "10", "20", "30", "40",
                           key="distance", value="RK_value") %>%
  mutate(distance = as.factor(as.numeric(distance)))
priors_RK_tidy <- priors_RK  %>%
  gather("2", "5", "10", "20","30", "40", key="distance", value="RK_value") %>%
  mutate(distance = as.factor(as.numeric(distance)))
  # filter(scenario=="random")
random_RK <- priors_RK_tidy %>% filter(scenario=="Random")
  
RK_plot <- ggplot(RKs_tidy, aes(distance, RK_value)) +
  geom_violin(fill="darkgreen", color="darkgreen", alpha=0.8, position=position_nudge(x=0.1), scale="width", width=0.5) +
  #geom_line(colour="grey", linetype="dashed", size=1.5, alpha=0.5) +
  #geom_errorbar(aes(ymin=mean_RK-sd_RK, ymax=mean_RK+sd_RK, width=0.1),
  #              position=position_dodge(width=0.5)) +
  ggtitle("Ripley's K")+
  ylab("K")+
  ylim(c(-0.5,20))+
  theme(text = element_text(size=12))+
  geom_violin(data=priors_RK_tidy, aes(x=distance, y=RK_value),
              fill="grey", color="grey", position=position_nudge(x=-0.1), alpha=0.8, scale="width", width=0.5)+
  geom_violin(data=random_RK, aes(x=distance, y=RK_value),
              fill="gold", alpha=0.8, colour="gold", scale="width", width=0.5)+
  geom_point(data=field_RK_tidy, aes(x=distance, y=value), fill="royalblue1", shape=21, size=3)+
  theme_bw()
  #geom_line(data=field_RK_tidy, aes(x=distance, y=value), colour='gold', size=1.5)
RK_plot


#network analysis
NAs_tidy <- NAs %>%
  gather(NA_stat, value, "internal loops":"transitivity") %>%
  group_by(scenario)

field_NA_tidy <- field_NA %>% gather(key="type_NA", value="NA_value",
                                     "internal loops":"transitivity")%>%
  group_by(type_NA)
NAs_tidy <- NAs %>% gather("internal loops":"transitivity", key="type_NA", value="NA_value")
priors_NA_tidy <- priors_NA  %>%
  gather("internal loops":"transitivity", key="type_NA", value="NA_value")
random_NA <- priors_NA_tidy %>% filter(scenario=="Random")

NA_plot <- ggplot(NAs_tidy, aes(x=type_NA,y=NA_value))+
  geom_violin(fill="darkgreen", color="darkgreen", alpha=0.8, position=position_nudge(x=0.1)) +
  facet_wrap(~type_NA, scales='free', nrow=1)+
  geom_violin(data=priors_NA_tidy, aes(x=type_NA, y=NA_value),
              fill="grey", color="grey", position=position_nudge(-0.1), alpha=0.8)+
  geom_hline(data=field_NA_tidy, aes(yintercept=NA_value), color='royalblue1', linetype="dashed", size=1)+
   geom_violin(data=random_NA, aes(x=type_NA, y=NA_value),
              fill="gold", alpha=0.2, colour="gold")+
  ylab("value")+
  ggtitle("Network metrics")+
  theme_bw()+
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        text = element_text(size=12), axis.title.x=element_blank())
NA_plot
  
grid.arrange(RK_plot, NA_plot, ncol=1, nrow=2, layout_matrix=rbind(c(1,1),c(2,2)), heights=c(1,0.75))

```


##Plot of priors and posterios of parameters and summary stats
Priors are in pink, posteriors in black.  

```{r}
#ALL
#priors and posteriors of the parameters
params_sel <- params %>% select(-pf, -scenario) %>% mutate_all(as.numeric)
params_prior <- params_sel %>% gather()
params_post <- params_sel[index_ss,] %>% gather()

#priors and posteriors of the summary stats
stats_prior <- sumstats_sel %>% select(-accepted) %>% gather()
stats_post <- sumstats_sel[index_ss,] %>% select(-accepted) %>% gather()

#parameters
#plot params all
ggplot(data=params_post, aes(x=value)) +
  geom_histogram(aes(y=..count../sum(..count..))) + facet_wrap(~key, scales='free')+
  geom_histogram(data=params_prior, aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5)

#summary stats
#plot summary stats all

ggplot(data=stats_post, aes(x=value)) +
  geom_histogram(aes(x=value, y=..count../sum(..count..))) + facet_wrap(~key, scales='free') +
  geom_vline(data=gather(ss_field_sel[1,]), aes(xintercept=value), color='green', linetype="dashed", size=3)+
  theme(text = element_text(size=12))+
  geom_histogram(data=stats_prior, aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5)
```

##Analysis for submodels separately
To narrow down parameters (this is the prior predictive check), and make sure not to exclude parameter values that might be important for one of the three submodels.

###UNIFORM

```{r}
#select uniform and random scenarios
sumstats_sel_uniform <- sumstats_sel_scen %>% filter(scenario=="UNIFORM" | scenario=='Random')
#get indexes of the data
indexes_uniform <- as.vector(unlist(sumstats_sel_uniform$scenario))
str(indexes_uniform)
sumstats_sel_uniform <- sumstats_sel_uniform  %>% select(-scenario)
#standardise data, according to simulations
sumstats_sel_uniform_sd <- scale(sumstats_sel_uniform, center=summarise_all(sumstats_sel_uniform, mean), scale=summarise_all(sumstats_sel_uniform, sd))

#also standardise field data, according to simulations
ss_field_sel_uniform_sd <- scale(ss_field_sel[1,], center=summarise_all(sumstats_sel_uniform, mean), scale=summarise_all(sumstats_sel_uniform, sd))

#get params for uniform
params_uniform <- params %>% filter(scenario=="UNIFORM" | scenario=='Random') %>% select(-pf, -scenario) %>% mutate_all(as.numeric)
#Calculate weights
pcor_uniform <- cor(sumstats_sel_uniform)
n_st_uniform <- nrow(pcor_uniform)

#take the mean of the pearson's correlation coefficients per variable
#then, mutate to get mean without the 1 from the diagonal of the matrix (which is x*n/(n-1) + 1/n)
mean_pcor_uniform <- as.data.frame(pcor_uniform) %>% mutate_all(abs) %>% summarise_all(mean) %>% mutate_all(function(x) {x*n_st_uniform/(n_st_uniform-1) - 1/n_st_uniform})
weights_uniform <- mean_pcor_uniform %>% mutate_all(function(x) {1-x}) #maybe to the power of 2?

#ABC analysis 100/50,000 = 0.002
abc_uniform <- create.abcSel(target=ss_field_sel_v, index=indexes_uniform,
                            results=as.matrix(sumstats_sel_uniform),
                            number.accept = 100, weights=as.numeric(weights_uniform[1,]))
summary(abc_uniform)

sumstats_sel_uniform$accepted <- abc_uniform$accepted
index_ss_uniform <- as.numeric(which(sumstats_sel_uniform$accepted==T))
stats_ss_uniform <- sumstats_sel_uniform[index_ss_uniform,]

#priors and posteriors of the parameters
params_prior_uniform <- params_uniform %>% gather()
params_post_uniform <- params_uniform[index_ss_uniform,] %>% gather()

#priors and posteriors of the summary stats
stats_prior_uniform <- sumstats_sel_uniform %>% select(-accepted) %>% gather()
stats_post_uniform <- sumstats_sel_uniform[index_ss_uniform,] %>% select(-accepted) %>% gather()

####Parameters plot
ggplot(data=params_post_uniform, aes(x=value)) +
  geom_histogram(aes(y=..count../sum(..count..))) + facet_wrap(~key, scales='free')+
  geom_histogram(data=params_prior_uniform, aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5)

####Summary stats plot
ggplot(data=stats_post_uniform, aes(x=value)) +
  geom_histogram(aes(x=value, y=..count../sum(..count..))) + facet_wrap(~key, scales='free') +
  geom_vline(data=gather(ss_field_sel[1,]), aes(xintercept=value), color='green', linetype="dashed", size=3)+
  theme(text = element_text(size=12))+
  geom_histogram(data=stats_prior, aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5)
```

###FIXED

```{r}
#select fixed and random scenarios
sumstats_sel_fixed <- sumstats_sel_scen %>% filter(scenario=="FIXED" | scenario=='Random')
#get indexes of the data
indexes_fixed <- as.vector(unlist(sumstats_sel_fixed$scenario))
str(indexes_fixed)
sumstats_sel_fixed <- sumstats_sel_fixed  %>% select(-scenario)
#standardise data, according to simulations
sumstats_sel_fixed_sd <- scale(sumstats_sel_fixed, center=summarise_all(sumstats_sel_fixed, mean), scale=summarise_all(sumstats_sel_fixed, sd))

#also standardise field data, according to simulations
ss_field_sel_fixed_sd <- scale(ss_field_sel[1,], center=summarise_all(sumstats_sel_fixed, mean), scale=summarise_all(sumstats_sel_fixed, sd))

#get params for fixed
params_fixed <- params %>% filter(scenario=="FIXED" | scenario=='Random') %>% select(-pf, -scenario) %>% mutate_all(as.numeric)
#Calculate weights
pcor_fixed <- cor(sumstats_sel_fixed)
n_st_fixed <- nrow(pcor_fixed)

#take the mean of the pearson's correlation coefficients per variable
#then, mutate to get mean without the 1 from the diagonal of the matrix (which is x*n/(n-1) + 1/n)
mean_pcor_fixed <- as.data.frame(pcor_fixed) %>% mutate_all(abs) %>% summarise_all(mean) %>% mutate_all(function(x) {x*n_st_fixed/(n_st_fixed-1) - 1/n_st_fixed})
weights_fixed <- mean_pcor_fixed %>% mutate_all(function(x) {1-x}) #maybe to the power of 2?

#ABC analysis 100/50,000 = 0.002
abc_fixed <- create.abcSel(target=ss_field_sel_v, index=indexes_fixed,
                            results=as.matrix(sumstats_sel_fixed),
                            number.accept = 100, weights=as.numeric(weights_fixed[1,]))
summary(abc_fixed)

sumstats_sel_fixed$accepted <- abc_fixed$accepted
index_ss_fixed <- as.numeric(which(sumstats_sel_fixed$accepted==T))
stats_ss_fixed <- sumstats_sel_fixed[index_ss_fixed,]

#priors and posteriors of the parameters
params_prior_fixed <- params_fixed %>% gather()
params_post_fixed <- params_fixed[index_ss_fixed,] %>% gather()

#priors and posteriors of the summary stats
stats_prior_fixed <- sumstats_sel_fixed %>% select(-accepted) %>% gather()
stats_post_fixed <- sumstats_sel_fixed[index_ss_fixed,] %>% select(-accepted) %>% gather()

####Parameters plot
ggplot(data=params_post_fixed, aes(x=value)) +
  geom_histogram(aes(y=..count../sum(..count..))) + facet_wrap(~key, scales='free')+
  geom_histogram(data=params_prior_fixed, aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5)

####Summary stats plot
ggplot(data=stats_post_fixed, aes(x=value)) +
  geom_histogram(aes(x=value, y=..count../sum(..count..))) + facet_wrap(~key, scales='free') +
  geom_vline(data=gather(ss_field_sel[1,]), aes(xintercept=value), color='green', linetype="dashed", size=3)+
  theme(text = element_text(size=12))+
  geom_histogram(data=stats_prior, aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5)
```


###FLEXIBLE

```{r}
#select flexible and random scenarios
sumstats_sel_flexible <- sumstats_sel_scen %>% filter(scenario=="FLEXIBLE" | scenario=='Random')
#get indexes of the data
indexes_flexible <- as.vector(unlist(sumstats_sel_flexible$scenario))
str(indexes_flexible)
sumstats_sel_flexible <- sumstats_sel_flexible  %>% select(-scenario)
#standardise data, according to simulations
sumstats_sel_flexible_sd <- scale(sumstats_sel_flexible, center=summarise_all(sumstats_sel_flexible, mean), scale=summarise_all(sumstats_sel_flexible, sd))

#also standardise field data, according to simulations
ss_field_sel_flexible_sd <- scale(ss_field_sel[1,], center=summarise_all(sumstats_sel_flexible, mean), scale=summarise_all(sumstats_sel_flexible, sd))

#get params for flexible
params_flexible <- params %>% filter(scenario=="FLEXIBLE" | scenario=='Random') %>% select(-pf, -scenario) %>% mutate_all(as.numeric)
#Calculate weights
pcor_flexible <- cor(sumstats_sel_flexible)
n_st_flexible <- nrow(pcor_flexible)

#take the mean of the pearson's correlation coefficients per variable
#then, mutate to get mean without the 1 from the diagonal of the matrix (which is x*n/(n-1) + 1/n)
mean_pcor_flexible <- as.data.frame(pcor_flexible) %>% mutate_all(abs) %>% summarise_all(mean) %>% mutate_all(function(x) {x*n_st_flexible/(n_st_flexible-1) - 1/n_st_flexible})
weights_flexible <- mean_pcor_flexible %>% mutate_all(function(x) {1-x}) #maybe to the power of 2?

#ABC analysis 100/50,000 = 0.002
abc_flexible <- create.abcSel(target=ss_field_sel_v, index=indexes_flexible,
                            results=as.matrix(sumstats_sel_flexible),
                            number.accept = 100, weights=as.numeric(weights_flexible[1,]))
summary(abc_flexible)

sumstats_sel_flexible$accepted <- abc_flexible$accepted
index_ss_flexible <- as.numeric(which(sumstats_sel_flexible$accepted==T))
stats_ss_flexible <- sumstats_sel_flexible[index_ss_flexible,]

#priors and posteriors of the parameters
params_prior_flexible <- params_flexible %>% gather()
params_post_flexible <- params_flexible[index_ss_flexible,] %>% gather()

#priors and posteriors of the summary stats
stats_prior_flexible <- sumstats_sel_flexible %>% select(-accepted) %>% gather()
stats_post_flexible <- sumstats_sel_flexible[index_ss_flexible,] %>% select(-accepted) %>% gather()

####Parameters plot
ggplot(data=params_post_flexible, aes(x=value)) +
  geom_histogram(aes(y=..count../sum(..count..))) + facet_wrap(~key, scales='free')+
  geom_histogram(data=params_prior_flexible, aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5)

####Summary stats plot
ggplot(data=stats_post_flexible, aes(x=value)) +
  geom_histogram(aes(x=value, y=..count../sum(..count..))) + facet_wrap(~key, scales='free') +
  geom_vline(data=gather(ss_field_sel[1,]), aes(xintercept=value), color='green', linetype="dashed", size=3)+
  theme(text = element_text(size=12))+
  geom_histogram(data=stats_prior, aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5)
```



```{r}
#plot link between nodes and parameters

#node ENV
ggplot(data=gather(params_sel[index_ss,], "param", "value", 5:8), aes(x=node_ENV, y=value)) +
  geom_point() + facet_wrap(~param, scales='free', nrow=1) +
  ggtitle("All models 100/200 000")+
  geom_hline(data = data.frame(yint=-3,param="param_mindens_ca"), aes(yintercept = yint), linetype = "dotted")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

ggplot(data=gather(params_uniform[index_ss_uniform,], "param", "value", 5:8), aes(x=node_ENV, y=value)) +
  geom_point() + facet_wrap(~param, scales='free', nrow=1) +
  ggtitle ("Uniform 100/50 000")+
  geom_hline(data = data.frame(yint=-3,param="param_mindens_ca"), aes(yintercept = yint), linetype = "dotted")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

ggplot(data=gather(params_fixed[index_ss_fixed,], "param", "value", 5:8), aes(x=node_ENV, y=value)) +
  geom_point() + facet_wrap(~param, scales='free', nrow=1) +
  ggtitle ("Fixed 100/50 000")+
  geom_hline(data = data.frame(yint=-3,param="param_mindens_ca"), aes(yintercept = yint), linetype = "dotted")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

ggplot(data=gather(params_flexible[index_ss_flexible,], "param", "value", 5:8), aes(x=node_ENV, y=value)) +
  geom_point() + facet_wrap(~param, scales='free', nrow=1) +
  ggtitle ("Flexible 100/50 000")+
  geom_hline(data = data.frame(yint=-3,param="param_mindens_ca"), aes(yintercept = yint), linetype = "dotted")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

#node CA
ggplot(data=gather(params_sel[index_ss,], "param", "value", 5:8), aes(x=node_CA, y=value)) +
  geom_point() + facet_wrap(~param, scales='free', nrow=1) +
  ggtitle("All models 100/200 000") +
  geom_hline(data = data.frame(yint=-3,param="param_mindens_ca"), aes(yintercept = yint), linetype = "dotted")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

ggplot(data=gather(params_uniform[index_ss_uniform,], "param", "value", 5:8), aes(x=node_CA, y=value)) +
  geom_point() + facet_wrap(~param, scales='free', nrow=1) +
  ggtitle ("Uniform 100/50 000")+
  geom_hline(data = data.frame(yint=-3,param="param_mindens_ca"), aes(yintercept = yint), linetype = "dotted")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

ggplot(data=gather(params_fixed[index_ss_fixed,], "param", "value", 5:8), aes(x=node_CA, y=value)) +
  geom_point() + facet_wrap(~param, scales='free', nrow=1) +
  ggtitle ("Fixed 100/50 000")+
  geom_hline(data = data.frame(yint=-3,param="param_mindens_ca"), aes(yintercept = yint), linetype = "dotted")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

ggplot(data=gather(params_flexible[index_ss_flexible,], "param", "value", 5:8), aes(x=node_CA, y=value)) +
  geom_point() + facet_wrap(~param, scales='free', nrow=1) +
  ggtitle ("Flexible 100/50 000")+
  geom_hline(data = data.frame(yint=-3,param="param_mindens_ca"), aes(yintercept = yint), linetype = "dotted")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

#node LSF
ggplot(data=gather(params_sel[index_ss,], "param", "value", 5:8), aes(x=node_LSF, y=value)) +
  geom_point() + facet_wrap(~param, scales='free', nrow=1) +
  ggtitle("All models 100/200 000")+
  geom_hline(data = data.frame(yint=-3,param="param_mindens_ca"), aes(yintercept = yint), linetype = "dotted")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

ggplot(data=gather(params_uniform[index_ss_uniform,], "param", "value", 5:8), aes(x=node_LSF, y=value)) +
  geom_point() + facet_wrap(~param, scales='free', nrow=1) +
  ggtitle ("Uniform 100/50 000")+
  geom_hline(data = data.frame(yint=-3,param="param_mindens_ca"), aes(yintercept = yint), linetype = "dotted")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

ggplot(data=gather(params_fixed[index_ss_fixed,], "param", "value", 5:8), aes(x=node_LSF, y=value)) +
  geom_point() + facet_wrap(~param, scales='free', nrow=1) +
  ggtitle ("Fixed 100/50 000")+
  geom_hline(data = data.frame(yint=-3,param="param_mindens_ca"), aes(yintercept = yint), linetype = "dotted")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

ggplot(data=gather(params_flexible[index_ss_flexible,], "param", "value", 5:8), aes(x=node_LSF, y=value)) +
  geom_point() + facet_wrap(~param, scales='free', nrow=1) +
  ggtitle ("Flexible 100/50 000")+
  geom_hline(data = data.frame(yint=-3,param="param_mindens_ca"), aes(yintercept = yint), linetype = "dotted")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")


#covariation for other parameters (not the nodes)
##sigmaca vs rangeca
a <- ggplot(data=params_sel[index_ss,], aes(x=param_sigma_ca, y=range_ca)) +
  geom_point() + #facet_wrap(~param, scales='free') +
  ggtitle("All models 100/200 000") +
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

b <- ggplot(data=params_uniform[index_ss_uniform,], aes(x=param_sigma_ca, y=range_ca)) +
  geom_point() + #facet_wrap(~param, scales='free') +
  ggtitle ("Uniform 100/50 000")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

c <- ggplot(data=params_fixed[index_ss_fixed,], aes(x=param_sigma_ca, y=range_ca)) +
  geom_point() + #facet_wrap(~param, scales='free') +
  ggtitle ("Fixed 100/50 000")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

d <- ggplot(data=params_flexible[index_ss_flexible,], aes(x=param_sigma_ca, y=range_ca)) +
  geom_point() + #facet_wrap(~param, scales='free') +
  ggtitle ("Flexible 100/50 000")+
  geom_hline(data = data.frame(yint=7,param="range_ca"), aes(yintercept = yint), linetype = "dotted")

grid.arrange(a,b,c,d, nrow=1)

##mindensca vs sigmaca
ggplot(data=params_sel[index_ss,], aes(x=param_mindens_ca, y=param_sigma_ca)) +
  geom_point() + #facet_wrap(~param, scales='free') +
  ggtitle("All models 100/200 000")+
  geom_vline(data = data.frame(xint=-3,param="param_mindens_ca"), aes(xintercept = xint), linetype = "dotted")

ggplot(data=params_uniform[index_ss_uniform,], aes(x=param_mindens_ca, y=param_sigma_ca)) +
  geom_point() + #facet_wrap(~param, scales='free') +
  ggtitle ("Uniform 100/50 000")+
  geom_vline(data = data.frame(xint=-3,param="param_mindens_ca"), aes(xintercept = xint), linetype = "dotted")

ggplot(data=params_fixed[index_ss_fixed,], aes(x=param_mindens_ca, y=param_sigma_ca)) +
  geom_point() + #facet_wrap(~param, scales='free') +
  ggtitle ("Fixed 100/50 000")+
  geom_vline(data = data.frame(xint=-3,param="param_mindens_ca"), aes(xintercept = xint), linetype = "dotted")

ggplot(data=params_flexible[index_ss_flexible,], aes(x=param_mindens_ca, y=param_sigma_ca)) +
  geom_point() + #facet_wrap(~param, scales='free') +
  ggtitle ("Flexible 100/50 000")+
  geom_vline(data = data.frame(xint=-3,param="param_mindens_ca"), aes(xintercept = xint), linetype = "dotted")
```

```{r}
pairs.panels(params_sel[index_ss,], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = FALSE
             )
pairs.panels(params_uniform[index_ss_uniform,], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE, # show density plots
             ellipses = FALSE
             )

pairs.panels(params_fixed[index_ss_fixed,], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE, # show density plots
             ellipses = FALSE
             )

pairs.panels(params_flexible[index_ss_flexible,], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE, # show density plots
             ellipses = FALSE
             )
```

