---
title: "ABC_Bembix_narrowpriors"
author: ""
date: "March 2021"
output: html_document
---

 
```{r, message=FALSE, warning=FALSE}
library(abc)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(gridExtra)
library(data.table)
library(psych)

source("H_modelselectionABC.R") #file with custom made ABC analysis (for doing the weighted model selection)

#loading the data: summary stats, params scenarios, field data summary stats
##summary stats
getwd()

multimerge = function(path, pattern){
  filenames=list.files(path=path, pattern=pattern, full.names=T)
  rbindlist(lapply(filenames, function(x){read.table(x, header=T)}))
}

sumstats <- multimerge(path="./Data outputs/Official runs", pattern="Summary_stats*")
nrow(sumstats)

colnames(sumstats) <- c("scenario", "file title", 'RK2', 'RK5', 'RK10', 'RK15', 'RK20', 'RK30', 'RK40',
                      "T0", "T2", "T4",  "T6", "T8", "T10", "T12", "T14",
                      "NA_internal_loops", "NA_all_loops","NA_dens_undirected", "NA_dens_directed",
                      "NA_reciproc","NA_transitivity_und","NA_transitivity_dir")
sumstats <- sumstats %>% select(-T0, -T2, -T4, -T6, -T8, -T10, -T12, -T14)

#drop the NA's, which stems from some runs that don't have interchange between clusters
#sumstats <- sumstats %>% drop_na()
ind_na <- which(is.na(sumstats), arr.ind=TRUE)[,1]
#get sumstats only for non-NA
sumstats <- sumstats %>% slice(-ind_na)
nrow(sumstats)

#Take relative Ripley's K to complete spatial randomness (CRS) (K-CRS)/CRS; CRS=pi*r^2
sumstats <- sumstats %>% mutate(RK2rel = ( RK2 - pi*2^2 )/( pi*2^2 ),
                                RK5rel = ( RK5 - pi*5^2 )/( pi*5^2 ),
                                RK10rel = ( RK10 - pi*10^2 )/( pi*10^2 ),
                                RK15rel = ( RK15 - pi*15^2 )/( pi*15^2 ),
                                RK20rel = ( RK20 - pi*20^2 )/( pi*20^2 ),
                                RK30rel = ( RK30 - pi*30^2 )/( pi*30^2 ),
                                RK40rel = ( RK40 - pi*40^2 )/( pi*40^2 ))

#parameters
params <- multimerge(path="./Data outputs/Official runs", pattern="Param*")
nrow(params)

head(params)
summary(params)

#values for NA's
params_na <- params %>% slice(ind_na) #--> small values of sigma_lsf mainly

#get only params for non-NA
params <- params %>% slice(-ind_na)
```


###Field data
Master thesis data  
```{r}
##field data summary stats
ss_field <- read.table("./Field data analyses/Summary_stats_field.txt", header=T)
ss_field$X0 <- NULL
ss_field$scenario <- NULL
ss_field[] <- lapply(ss_field[], as.numeric)
colnames(ss_field) <- c("mean_RK", "sd_RK", 'RK2', 'RK5', 'RK10', 'RK15', 'RK20', 'RK30', 'RK40',
                           "NA_internal_loops", "NA_all_loops","NA_dens_undirected", "NA_dens_directed",
                           "NA_reciproc","NA_transitivity_und","NA_transitivity_dir")
ss_field$mean_RK <- NULL
ss_field$sd_RK <- NULL

#Take relative Ripley's K to complete spatial randomness (CRS) (K-CRS)/CRS; CRS=pi*r^2
ss_field <- ss_field %>% mutate(RK2rel = ( RK2 - pi*2^2 )/( pi*2^2 ),
                                RK5rel = ( RK5 - pi*5^2 )/( pi*5^2 ),
                                RK10rel = ( RK10 - pi*10^2 )/( pi*10^2 ),
                                RK15rel = ( RK15 - pi*15^2 )/( pi*15^2 ),
                                RK20rel = ( RK20 - pi*20^2 )/( pi*20^2 ),
                                RK30rel = ( RK30 - pi*30^2 )/( pi*30^2 ),
                                RK40rel = ( RK40 - pi*40^2 )/( pi*40^2 ))

#make a 1D vector of it
ss_field_v <- as.numeric(ss_field[1,])
```


```{r}
#keep summary stats with scenario attached to df
sumstats_wscen <- sumstats
#make sure the summary stats are numeric values (scenario will be replaced with 1, 2, 3)
sumstats[] <- lapply(sumstats[], as.numeric) #make sure all are numeric values

sumstats$scenario <- NULL #take scenario out of summary stats, to make calculations easier
sumstats$`file title` <- NULL
```


####Correlation between summary stats
Within the Ripley's K, there is high correlation. But less for the high values. We should select distances that are further apart from each other. F.i. RK2, RK10, RK20, RK30, RK40.  
Within Network metrics, there is some correlation. Especially for the same metric for undirected and directed networks. These will be dropped.  
For the network metrics and Ripley's K, there is some structure. But this is expected with such high level network metrics. Variation seems to be high enough.  

```{r}
#make selection
sumstats_sel <- sumstats %>% select("RK2rel","RK5rel", "RK10rel", "RK20rel", "RK30rel", "RK40rel",
                                 "NA_internal_loops", "NA_all_loops", "NA_dens_directed",
                                 "NA_reciproc", "NA_transitivity_dir")

ss_field_sel <- ss_field %>% select("RK2rel","RK5rel", "RK10rel", "RK20rel", "RK30rel", "RK40rel",
                                 "NA_internal_loops", "NA_all_loops", "NA_dens_directed",
                                 "NA_reciproc", "NA_transitivity_dir")

#standardise data, according to simulations
sumstats_sel_sd <- scale(sumstats_sel, center=summarise_all(sumstats_sel, mean), scale=summarise_all(sumstats_sel, sd))
#also standardise field data, according to simulaitons
ss_field_sel_sd <- scale(ss_field_sel[1,], center=summarise_all(sumstats_sel, mean), scale=summarise_all(sumstats_sel, sd))


#Within Ripley's K
pairs(sumstats_sel_sd[sample(nrow(sumstats_sel_sd), 1000),c(1:6)])
#within network metrics
pairs(sumstats_sel_sd[sample(nrow(sumstats_sel_sd), 1000), c(7:11)])

#make field data vector
ss_field_sel_v <- as.numeric(ss_field_sel_sd[1,])
#make a sumstats selected with scenario
sumstats_sel_scen <- sumstats_sel #not with standardised data
sumstats_sel_scen <- bind_cols(sumstats_sel_scen, "scenario"=sumstats_wscen$scenario)
head(sumstats_sel_scen)
```

###Indexes
A vector has to be created with the names in of the models corresponding to the different runs.  

```{r}
#indexes -> which row corresponds with which model
indexes <- as.vector(unlist(sumstats_wscen[,1]))
str(indexes)
```



####Making Pearson's correlation matrix

```{r}
#calculate pearson's correlation coefficients for all
pcor <- cor(sumstats_sel_sd)
#View(pcor)

n_st <- nrow(pcor)

#take the mean of the pearson's correlation coefficients per variable
#then, mutate to get mean without the 1 from the diagonal of the matrix (which is x*n/(n-1) + 1/n)
mean_pcor <- as.data.frame(pcor) %>% mutate_all(abs) %>% summarise_all(mean) %>% mutate_all(function(x) {x*n_st/(n_st-1) - 1/n_st})
weights <- mean_pcor %>% mutate_all(function(x) {1-x}) #maybe to the power of 2?

weights
```


##Cross validation for model selection

runs quite long, uncomment to run

```{r eval=FALSE, include=FALSE}
# cv_bembix <- abc::cv4postpr(index=indexes, sumstat=sumstats_sel_sd, nval=50, tols=0.001003272, method='rejection')
# summary(cv_bembix)
# plot(cv_bembix, names.arg=c("Fixed", "Flexible", "Random", "Uniform"), probs=TRUE)
# plot(cv_bembix, names.arg=c("Fixed", "Flexible", "Random", "Uniform"), probs=FALSE)
```

##Model selection
'postpr' function from the package abc is used to look which model is mostly selected (when summary stats lay closest to field data sumstats). 

```{r}
# modsel_bembix <- postpr(target=ss_field_sel_v, index=indexes, sumstat=sumstats_sel_sd,tol=0.001,method="rejection")
# summary(modsel_bembix)

modsel_1000 <- create.abcSel(target=ss_field_sel_v, indexes=indexes,
                             results=as.matrix(sumstats_sel_sd),
                             number.accept = 1000, weights=as.numeric(weights[1,]))
# modsel_1000 <- create.abcSel(target=ss_field_sel_v, indexes=indexes, results=as.matrix(sumstats_sel), rate=0.01, weights=rep(1, ncol(sumstats_sel)))
summary(modsel_1000)

modsel_100 <- create.abcSel(target=ss_field_sel_v, index=indexes,
                            results=as.matrix(sumstats_sel_sd),
                            number.accept = 100, weights=as.numeric(weights[1,]))
summary(modsel_100)

#500?
modsel_500 <- create.abcSel(target=ss_field_sel_v, index=indexes,
                            results=as.matrix(sumstats_sel_sd),
                            number.accept = 500, weights=as.numeric(weights[1,]))
summary(modsel_500)

#10,000
modsel_ten1000 <- create.abcSel(target=ss_field_sel_v, index=indexes,
                            results=as.matrix(sumstats_sel_sd),
                            number.accept = 10000, weights=as.numeric(weights[1,]))
summary(modsel_ten1000)

#source("D:/fbatslee/OneDrive - UGent/Conferences/BES/ABCWorkshop/Code/R/ModelSelection.R")
```

##Summary stats of models that were accepted

```{r, message=FALSE, warning=FALSE}
#index_ss <- as.numeric(rownames(modsel_bembix$ss))
sumstats_sel$accepted <- modsel_1000$accepted
index_ss <- as.numeric(which(sumstats_sel$accepted==T))
stats_ss <- sumstats_sel[index_ss,]

# ggplot(data=gather(stats_ss), aes(x=value)) +
#   geom_histogram(aes(x=value, y=..count../sum(..count..))) + facet_wrap(~key, scales='free') +
#   geom_vline(data=gather(ss_field_sel[1,]), aes(xintercept=value), color='green', linetype="dashed", size=3)+
#   theme(text = element_text(size=12))+
#   geom_histogram(data=gather(sumstats_sel), aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5) +
#   theme_bw()
```

###Plot of the accepted summary stats with field data

Ripley's K: spatial clustering is quite well formed.  
Network analysis are produced quite well.  
```{r}
#get ripley's K and network analysis values for the accepted models
stats_ss_w_scen <- sumstats_sel_scen[index_ss,]
RKs <- stats_ss_w_scen %>% select("scenario","RK2rel", "RK5rel", "RK10rel", "RK20rel","RK30rel", "RK40rel")
NAs <- stats_ss_w_scen %>% select("scenario", "NA_internal_loops", "NA_all_loops",
                                  "NA_dens_directed","NA_reciproc", "NA_transitivity_dir")

# str(RKs)
# str(NAs)

#get field summary stats
field_RK <- ss_field_sel %>%
  select("RK2rel","RK5rel", "RK10rel", "RK20rel","RK30rel", "RK40rel") %>%
  filter(row_number()==1)
field_RK$scenario <- 'field data'

field_NA <- ss_field_sel %>% select("NA_internal_loops", "NA_all_loops",
                                  "NA_dens_directed","NA_reciproc", "NA_transitivity_dir") %>%
  filter(row_number()==1)

#priors

priors_RK <- sumstats_sel_scen  %>%
  select("scenario", "RK2rel","RK5rel", "RK10rel", "RK20rel", "RK30rel", "RK40rel")
priors_NA <- sumstats_sel_scen %>% select("scenario","NA_internal_loops", "NA_all_loops",
                                  "NA_dens_directed","NA_reciproc", "NA_transitivity_dir")

#change colnames to easily plot the functions correctly
colnames(RKs) <- c("scenario",'2', '5', '10', '20', '30', '40')
colnames(field_RK) <- c('2', '5', '10', '20', '30', '40',"scenario")
colnames(priors_RK) <- c('scenario','2', '5', '10', '20','30', '40')


colnames(NAs) <- c("scenario",'internal loops','all loops','density',
                   'reciprocity', 'transitivity')
colnames(field_NA) <- c('internal loops','all loops','density',
                   'reciprocity', 'transitivity')
colnames(priors_NA) <- c("scenario",'internal loops','all loops','density',
                   'reciprocity', 'transitivity')

#give the scenario's a name again (instead of a number)  
# RKs$scenario <- modsel_100$values
# NAs$scenario <- modsel_100$values

#Ripley's K graphs
# RKs_tidy <- RKs %>%
#   gather(distance, value, '2':'40') %>%
#   mutate(distance = as.factor(as.numeric(distance))) %>%
#   group_by(scenario, distance)
# RKs_summary <- RKs_tidy %>% summarise(mean_RK = mean(value), sd_RK = sd(value))
# 
field_RK_tidy <- field_RK %>% gather(distance, value, '2':'40') %>%
  mutate(distance = as.factor(as.numeric(distance)))
RKs_tidy <- RKs %>% gather("2","5", "10", "20", "30", "40",
                           key="distance", value="RK_value") %>%
  mutate(distance = as.factor(as.numeric(distance)))
priors_RK_tidy <- priors_RK  %>%
  gather("2", "5", "10", "20","30", "40", key="distance", value="RK_value") %>%
  mutate(distance = as.factor(as.numeric(distance)))
  # filter(scenario=="random")
random_RK <- priors_RK_tidy %>% filter(scenario=="Random")
  
RK_plot <- ggplot(RKs_tidy, aes(distance, RK_value)) +
  geom_violin(fill="darkgreen", color="darkgreen", alpha=0.8, position=position_nudge(x=0.1), scale="width", width=0.5) +
  #geom_jitter(color="black", size=0.4, alpha=0.9) +
  #geom_line(colour="grey", linetype="dashed", size=1.5, alpha=0.5) +
  #geom_errorbar(aes(ymin=mean_RK-sd_RK, ymax=mean_RK+sd_RK, width=0.1),
  #              position=position_dodge(width=0.5)) +
  ggtitle("Ripley's K")+
  ylab("(K-CSR)/CSR")+
  ylim(c(-0.5,20))+
  theme(text = element_text(size=12))+
  geom_violin(data=priors_RK_tidy, aes(x=distance, y=RK_value),
              fill="grey", color="grey", position=position_nudge(x=-0.1), alpha=0.8, scale="width", width=0.5)+
  geom_violin(data=random_RK, aes(x=distance, y=RK_value),
              fill="gold", alpha=0.8, colour="gold", scale="width", width=0.5)+
  geom_point(data=field_RK_tidy, aes(x=distance, y=value), fill="royalblue1", shape=21, size=3)+
  theme_bw()
  #geom_line(data=field_RK_tidy, aes(x=distance, y=value), colour='gold', size=1.5)
#RK_plot


#network analysis
NAs_tidy <- NAs %>%
  gather(NA_stat, value, "internal loops":"transitivity") %>%
  group_by(scenario)

field_NA_tidy <- field_NA %>% gather(key="type_NA", value="NA_value",
                                     "internal loops":"transitivity")%>%
  group_by(type_NA)
NAs_tidy <- NAs %>% gather("internal loops":"transitivity", key="type_NA", value="NA_value")
priors_NA_tidy <- priors_NA  %>%
  gather("internal loops":"transitivity", key="type_NA", value="NA_value")
random_NA <- priors_NA_tidy %>% filter(scenario=="Random")

NA_plot <- ggplot(NAs_tidy, aes(x=type_NA,y=NA_value))+
  geom_violin(fill="darkgreen", color="darkgreen", alpha=0.8, position=position_nudge(x=0.1)) +
  #geom_jitter(color="black", size=0.4, alpha=0.9) +
  facet_wrap(~type_NA, scales='free', nrow=1)+
  geom_violin(data=priors_NA_tidy, aes(x=type_NA, y=NA_value),
              fill="grey", color="grey", position=position_nudge(-0.1), alpha=0.8)+
  geom_hline(data=field_NA_tidy, aes(yintercept=NA_value), color='royalblue1', linetype="dashed", size=1)+
   geom_violin(data=random_NA, aes(x=type_NA, y=NA_value),
              fill="gold", alpha=0.2, colour="gold")+
  ylab("value")+
  ggtitle("Network metrics")+
  theme_bw()+
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        text = element_text(size=12), axis.title.x=element_blank())
#NA_plot
  
grid.arrange(RK_plot, NA_plot, ncol=1, nrow=2, layout_matrix=rbind(c(1,1),c(2,2)), heights=c(1,0.75))

```


##Plot of priors and posterios of parameters and summary stats
Priors are in pink, posteriors in black.  

```{r}
#priors and posteriors of the parameters
params_sel <- params %>% select(-pf, -scenario) %>% mutate_all(as.numeric)
params_prior <- params_sel %>% gather()
params_post <- params_sel[index_ss,] %>% gather()

params_post_scen <- params[index_ss,] %>% select(-pf) %>% mutate_all(as.numeric) %>%  gather(key="key", value="value",-scenario)

#priors and posteriors of the summary stats
stats_prior <- sumstats_sel %>% select(-accepted) %>% gather()
stats_post <- sumstats_sel[index_ss,] %>% select(-accepted) %>% gather()

####PARAMETERS
#plot params all
params_postprior <- ggplot(data=
         transform(params_post_scen,
                   key=factor(key, levels=c("node_ENV", "node_LSF", "node_CA","beh_excl",
                                            "sigma_lsf", "range_ca", "param_mindens_ca", "param_sigma_ca"
                                            ))),
       aes(x=value)) +
  stat_bin(aes(y=8*..count../sum(..count..)), boundary=0) + facet_wrap(~key, scales='free', nrow=2)+ #/sum(..count..)
  stat_bin(data=
             transform(params_prior, key=factor(key, levels=c("node_ENV", "node_LSF", "node_CA","beh_excl",
                                            "sigma_lsf", "range_ca", "param_mindens_ca", "param_sigma_ca"))),
           aes(x=value, y=8*..count../sum(..count..)), boundary=0,
           fill="#FF6666", alpha=0.5) +
  xlab("parameter value")+
  ylab("frequency (%)")+
  scale_y_continuous(labels=scales::percent)+
  theme_bw()
params_postprior


#####SUMMARY STATISTICS
#plot summary stats all

ggplot(data=stats_post, aes(x=value)) +
  geom_histogram(aes(x=value, y=..count../sum(..count..))) + facet_wrap(~key, scales='free') +
  geom_vline(data=gather(ss_field_sel[1,]), aes(xintercept=value), color='green', linetype="dashed", size=3)+
  theme(text = element_text(size=12))+
  geom_histogram(data=stats_prior, aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5)

##
pairs.panels(params_sel[index_ss,], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE, # show density plots
             ellipses = FALSE
             )
```

##Visualisation of response functions

```{r}
#posteriors of the parameters
posteriors <- params[index_ss,] %>% tibble::rownames_to_column("ID")


sigm_function <- function(mindens, sigmaca, densities){
  probs <- 1/(1+exp(-(mindens + sigmaca*densities)))
  return(probs)
}

####conspecific attraction####
#for parameter sigma_ca
#make density-vector
dens <- seq(from=0, to=30, by=0.05)
length(dens)

quantiles_sigma <- data.frame(value=quantile(posteriors$param_sigma_ca, c(0.2, 0.5, 0.8))) %>% bind_cols(data.frame(quant=c('20%', '50%', '80%')))
combi_sigma <- crossing(quantiles_sigma, dens)
#join them together to get all possible combinations of posteriors and density-vector
combi_full_sigma <- combi_sigma %>% mutate(resp = sigm_function(mindens=median(posteriors$param_mindens_ca),
                                                         sigmaca =value,
                                                         densities=dens)) %>%
  mutate(dens_rel = dens/(pi*median(posteriors$range_ca)^2))

#plot of response function
resp_ca_sigma <- ggplot(data=combi_full_sigma, aes(x=dens_rel, y=resp)) +
  geom_line(aes(linetype=quant, col=quant))+
  scale_linetype_manual(values=c("twodash", "solid", "dashed"))+
  scale_color_manual(values=c("blue", "black", "blue"))+
  xlab("density (# nest / m²)") + ylab("probability") + labs(linetype="Quantile", col="Quantile")+
  xlim(0,2.5)+
  theme_bw()+ theme(legend.position = c(0.75, 0.3))

#for parameters min_dens_ca
quantiles_mindens <- data.frame(value=quantile(posteriors$param_mindens_ca, c(0.2, 0.5, 0.8))) %>% bind_cols(data.frame(quant=c('20%', '50%', '80%')))
combi_mindens <- crossing(quantiles_mindens, dens)
combi_full_mindens <- combi_mindens %>% mutate(resp = sigm_function(mindens=value,
                                                         sigmaca =median(posteriors$param_sigma_ca),
                                                         densities=dens)) %>%
  mutate(dens_rel = dens/(pi*median(posteriors$range_ca)^2))
resp_ca_mindens <- ggplot(data=combi_full_mindens, aes(x=dens_rel, y=resp)) +
  geom_line(aes(linetype=quant, col=quant))+
  scale_linetype_manual(values=c("twodash", "solid", "dashed"))+
  scale_color_manual(values=c("blue", "black", "blue"))+
  xlab("density (# nests / m²)") + ylab("probability") + labs(linetype="Quantile", col="Quantile")+
  xlim(0,2.5)+
  theme_bw()+ theme(legend.position = c(0.75, 0.3))

#for parameter range_ca
quantiles_range_ca <- data.frame(value=quantile(posteriors$range_ca, c(0.2, 0.5, 0.8))) %>% bind_cols(data.frame(quant=c('20%', '50%', '80%')))
combi_range <- crossing(quantiles_range_ca, dens)

combi_full_range <- combi_range %>% mutate(resp = sigm_function(mindens=median(posteriors$param_mindens_ca),
                                                         sigmaca =median(posteriors$param_sigma_ca),
                                                         densities=dens)) %>%
  mutate(dens_rel = dens/(pi*value^2))

resp_ca_range <- ggplot(data=combi_full_range, aes(x=dens_rel, y=resp)) +
  geom_line(aes(linetype=quant, col=quant))+
  scale_linetype_manual(values=c("twodash", "solid", "dashed"))+
  scale_color_manual(values=c("blue", "black", "blue"))+
  xlab("density (# nests / m²)") + ylab("probability") + labs(linetype="Quantile", col="Quantile")+
  xlim(0,2.5)+
  theme_bw() + theme(legend.position = c(0.75, 0.3))

###Local site fidelity####
dist <- seq(from=0, to=10, by=0.05)
length(dist)
quantiles <- data.frame(value=quantile(posteriors$sigma_lsf, c(0.2, 0.5, 0.8))) %>% bind_cols(data.frame(quant=c('20%', '50%', '80%')))
#combine them with ID's of the posteriors
combi_lsf <- crossing(quantiles, dist)

gauss_function <- function(sigmalsf, distances){
  probs <- (1/(sigmalsf*sqrt(2*pi)))*exp((-distances^2)/(2*sigmalsf^2))
  return(probs)
}

combi_full_lsf <- combi_lsf %>% mutate(resp = gauss_function(sigmalsf=value, distances=dist))

resp_lsf <- ggplot(data=combi_full_lsf , aes(x=dist, y=resp)) +
  geom_line(aes(linetype=quant, col=quant))+#scale_colour_brewer(palette = "Spectral") +
  #xlim(0,5) +
  scale_linetype_manual(values=c("twodash", "solid", "dashed"))+
  scale_color_manual(values=c("blue", "black", "blue"))+
  xlab("distance (m)") + ylab("probability") + labs(linetype="Quantile", col="Quantile")+
  theme_bw()+ theme(legend.position = c(0.7, 0.7))



grid.arrange(resp_lsf, resp_ca_range, resp_ca_mindens, resp_ca_sigma, nrow=1)
```

